A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2020-05-13 manuscript.Rmd
added thanks note to GitHub contributors

authornote: |
   We are grateful to Julia Delius for her helpful assistance in language and style editing. This paper is fully reproducible using the workflow described in this paper. All materials can be found on the accompanying GitHub repository: [https://github.com/aaronpeikert/reproducible-research/](https://github.com/aaronpeikert/reproducible-research/). This particular document was created from the commit with the hash `"`r system2("git", "rev-parse --short HEAD", stdout = TRUE)`"`. To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the GitHub repository of our paper at [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).

@@

While our approach was designed to scale well with the complexity of a computing-intense project, we realise that this flexibility may not be straightforward to integrate into a researcher's everyday workflow. There are R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed. The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr]. The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R centric approach, making it especially suited for projects only involving R, but can also handle external dependencies. The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker. The former is perfectly compatible with the described workflow and we recommend it to users  who are not comfortable with command-line use of Docker. `holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of Docker is required. `holepunch` is well suited for simple analysis with low computational demands, because binder's  memory and computing time is limited. There are several alternatives to Docker that manage dependencies on R packages. [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@renv] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies. Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint]. The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited. A particular noteworthy approach is provided by the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs, @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with git, package management with renv and dynamic document generation with R Markdown. We acknowledge that worcs is much easier to install and provides a one-click solution for the creation of reproducible projects. It achieves a high standard of reproducibility but does not guarantuee full computational reproducibility and is limited to dependency management within the R environment.

@@

The proposed workflow leverages various existing tools that are partly integrated into RStudio already. Parts of the proposed workflow have been integrated into stand-alone packages (such as worcs [@worcs], workflowr [@workflowr], holepunch  [@holepunch]) but they are either incomplete in the sense of our proposal or rely heavily on propietary service providers. We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to an even better integration of existing tools to allow reproducible workflows to become a standard in psychological research.

>>>>

authornote: |
   We are grateful to Julia Delius for her helpful assistance in language and style editing. We thank all contributors of our GitHub repository for helpful feedback on the paper. This paper is fully reproducible using the workflow described in this paper. All materials can be found on the accompanying GitHub repository: [https://github.com/aaronpeikert/reproducible-research/](https://github.com/aaronpeikert/reproducible-research/). This particular document was created from the commit with the hash `"`r system2("git", "rev-parse --short HEAD", stdout = TRUE)`"`. To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the GitHub repository of our paper at [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).

@@

While our approach was designed to scale well with the complexity of a computing-intense project, we realise that this flexibility may not be straightforward to integrate into a researcher's everyday workflow. There are R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed. The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr]. The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R centric approach, making it especially suited for projects only involving R, but can also handle external dependencies. The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker. The former is perfectly compatible with the described workflow and we recommend it to users  who are not comfortable with command-line use of Docker. `holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of Docker is required. `holepunch` is well suited for simple analysis with low computational demands, because binder's  memory and computing time is limited. There are several alternatives to Docker that manage dependencies on R packages. [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@renv] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies. Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint]. The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited. A particularly noteworthy approach is provided by the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs, @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with git, package management with renv and dynamic document generation with R Markdown. We acknowledge that worcs is much easier to install and provides a one-click solution for the creation of reproducible projects. It achieves a high standard of reproducibility but does not guarantuee full computational reproducibility and is limited to dependency management within the R environment.

@@

The proposed workflow leverages various existing tools that are partly integrated into RStudio already. Parts of the proposed workflow have been integrated into stand-alone packages (such as worcs [@worcs], workflowr [@workflowr], holepunch  [@holepunch]), which are particularly recommenable to beginners but they are either incomplete in the sense of our proposal or rely on proprietary service providers. We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to an even better integration of existing tools to allow reproducible workflows to become a standard in psychological research.
